\hydrogen{} and \carbon{} Network Strategies for the European Energy System

### 2024-09-06 report/submissions/journal/paper/paper.tex
finish addressing review comments

\documentclass[twocolumn]{article}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{cuted}  % Add the cuted package
\usepackage{caption}
\captionsetup{font=small}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{tabularx, multirow}
\usepackage{hyperref} % Add this line to enable hyperlinks
\usepackage{fancyhdr,lipsum}
\usepackage{subcaption}
\usepackage[shortcuts]{extdash}

@@

To investigate the competing transport dynamics between \carbon{} and \hydrogen{} networks, we contrast the following four scenarios:
%
\begin{itemize}
    \item[] \textit{Baseline:} Neither \carbon{} nor \hydrogen{} networks are constructed.
    \item[] \textit{\carbongrid{}:} Only the \carbon{} network is developed.
    \item[] \textit{\hydrogengrid{}:} Only the \hydrogen{} network is developed.
    \item[] \textit{Hybrid:} Both \carbon{} and \hydrogen{} networks are developed.
\end{itemize}

@@

\subsection*{Stricter emission targets increase \carbon{} network benefits}
\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.9\linewidth]{difference/comparison/emission-reduction-0.1/figures/90_nodes/cost_bar.png}
    \caption[short]{Net change in system cost when tightening the \carbon{} emission target from net-zero to net-negative 10\% of 1990s emissions. Each cost bar is split into contributions of the same technology groups as used in Fig.~\ref{fig:cost_bar}, except for carbon sequestration costs, which are now displayed separately. For all models, Direct Air Capture (DAC) contributes the most to additional \carbon{} removal, requiring further solar, wind and heat pump capacities for electricity and heat input.}
    \label{fig:net-negative_cost_bar}
\end{figure}

When applying a stricter \carbon{} emissions target and forcing the model to remove 460~Mt \carbon{} over one year (equivalent to 10\% of \carbon{} emissions in 1990) while sequestering maximally 660~Mt, investment in DAC and associated solar and wind power and heat supply covers a large share of additional system costs in all scenarios (see Fig.~\ref{fig:net-negative_cost_bar}). In the \hybridscenario{}, this share is at 80\% while new carbon capture at point sources only makes up 6.5\% of new system costs. This shows that the exploitation of carbon point sources is close to its feasible limit. The additional sequestration is primarily supplied by large DAC facilities near the shore. To a much smaller extent, carbon captured from bio-energetic sources in the inland, that under a net-zero emissions assumption supplies CU, is instead transported and sequestered (compare Fig.~\ref{fig:balance_maps_full_nn}). Some of the CU which in net-zero scenarios locates in Central Europe instead locates in Spain, where it uses carbon from new DAC facilities.

In the \carbon{} and \hydrogen{} scenarios, this slight shift in CU deployment results in a lower buildout of hydrogen pipelines, especially from the Iberian Peninsula to Central Europe (0.5~bn€/a). Nonetheless, 80\% of the hydrogen topology which is built under net-zero emissions is also built under the assumption of net-negative emissions. Under net-negative emissions the \carbon{} network also includes routes from Central and Eastern Europe to the shores (1.2~bn€/a). The benefits of \carbon{} transport and sequestration with tighter emission targets are also evident when comparing the other scenarios. The carbon network under the net-negative \carbonscenario{} is larger than under the net-zero \carbonscenario{} and transports more carbon from point sources to sequestration sites rather than to CU facilities. In the \hydrogenscenario{}, where all carbon for CU and CS must come from DAC, the hydrogen transport topology remains almost identical to the net-zero scenarios (for detailed cost contributions, refer to sections~\ref{sec:cost_comparison} and~\ref{sec:operation_nn} in the Appendix).

Overall, this leads to the net cost increases of 77~bn€/a in the \carbonscenario{}, by 81~bn€/a in the \hydrogenscenario{}, and by 78~bn€/a in the \hybridscenario{}. The resulting cost-benefit of the \hybridscenario{} in comparison to the other scenarios is 6~bn€/a.

@@

\subsection{Limitations}
\label{sec:limitations}
Despite the comprehensive nature of the model, some limitations affect the validity of the results, as well as important discussion points that need to be put into context.
In reality, energy demand and renewable supply can only be estimated for the next few hours/days. The model's perfect foresight may lead to non-reproducible technology dispatch which neglects uncertainties from long-term predictions and necessary reserve margins.
Demand and emissions from industrial clusters heavily drive the modeling results. Allowing the model to relocate these, incorporate flexibility measures or alter underlying processes, may lead to less dependency on \carbon{} and \hydrogen{} networks.
The transport of fuel (FT, gas, oil) between the regions is not limited which may overestimate their flexibility. However, the distributed CU systems, as enabled through the hydrogen network, align well with today's distributed locations for oil refineries, providing kerosene for aviation and naphtha for industry. The assumption to run Fischer-Tropsch facilities on baseload with at least 90\% of their nominal capacity is likely shaping the deployment of CU, underestimating potential important flexibilities.
The assumed cost projections on technologies are subject to uncertainties. As shown in \ref{sec:subsidy}, the effect of a 50\% cost reduction on \carbon{} pipelines alters total system costs to a small extent (below 1\%). However, a different set of costs on DAC and electrolysis may well impact the model results.
In combination with a net \carbon{} neutrality or reduction target, a relaxation of the global sequestration limit drives system costs down by introducing fossil fuels, as shown in~\cite{hofmannDesigningCO2Network2023}. Allowing the model to sequester onshore, would likely reduce the dominant co-location of CU with the point sources due to easier access to sequestration sites.
Finally, DAC deployment could be in parts replaced if larger sustainable biomass sources are assumed, as shown in ~\cite{lauerCrucialRoleBioenergy2023}.

>>>>

\documentclass[twocolumn]{article}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{cuted}  % Add the cuted package
\usepackage{caption}
\captionsetup{font=small}
\usepackage{algorithmic}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{tabularx, multirow}
\usepackage{hyperref} % Add this line to enable hyperlinks
\usepackage{fancyhdr,lipsum}
\usepackage{subcaption}
\usepackage[shortcuts]{extdash}

@@

To investigate the competing transport dynamics between \carbon{} and \hydrogen{} networks, we contrast the following four scenarios with different sets of networks present:
%

\begin{table}[ht!]
    \centering
    \begin{tabular}{c|c|c}
        \toprule
        \textbf{Scenario} & \textbf{\carbon{} Network} & \textbf{\hydrogen{} Network} \\
        \midrule
        Baseline & -- & -- \\
        \carbongrid{} & \checkmark & -- \\
        \hydrogengrid{} & -- & \checkmark \\
        Hybrid & \checkmark & \checkmark \\
        \bottomrule
    \end{tabular}
    \caption{Set of transportation networks in the developed scenarios.}
    \label{tab:scenarios}
\end{table}

@@

\subsection*{Network layouts are robust againg tighter emission targets}
\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.9\linewidth]{difference/comparison/emission-reduction-0.1/figures/90_nodes/cost_bar.png}
    \caption[short]{Net change in system cost when tightening the \carbon{} emission target from net-zero to net-negative 10\% of 1990s emissions. Each cost bar is split into contributions of the same technology groups as used in Fig.~\ref{fig:cost_bar}, except for carbon sequestration costs, which are now displayed separately. For all models, Direct Air Capture (DAC) contributes the most to additional \carbon{} removal, requiring further solar, wind and heat pump capacities for electricity and heat input.}
    \label{fig:net-negative_cost_bar}
\end{figure}

Applying a stricter \carbon{} emissions target and forcing the model to remove 460~Mt of \carbon{} in one year (equivalent to 10\% of 1990 \carbon{} emissions) while sequestering a maximum of 660~Mt, investment in DAC and associated solar and wind power and heat supply covers a large proportion of additional system costs in all scenarios (see Fig.~\ref{fig:net-negative_cost_bar}). In the \hybridscenario{}, this share is 78\%, while new carbon capture at point sources accounts for only 4\% of incremental system costs. This shows that point source carbon capture is closing the gap to its feasible limit. The additional sequestration is mainly provided by large DAC facilities near the coast. To a much lesser extent, carbon captured from inland bioenergy sources, which would provide CU under a net-zero assumption, is instead transported and sequestered (see Fig.~\ref{fig:balance_maps_full_nn}). Some of the CU located in Central Europe in net-zero scenarios is instead located in Spain, where it uses carbon from new DAC facilities.

These changes have little effect on the layout of the \carbon{} and \hydrogen{} networks.
The slight shift in CU deployment results in less hydrogen pipeline construction in some parts of the system. Nevertheless, 80\% of the hydrogen topology built under the net-zero emissions assumption is also built under the net-negative emissions assumption. The \carbon{} network partially extends routes from Central and Eastern Europe to the coasts. The relative benefits of \carbon{} transport and sequestration with tighter emissions targets are evident when compared with the other scenarios. The carbon network in the net-negative \carbonscenario{} is larger than in the net-zero \carbonscenario{} and transports more carbon from point sources to sequestration sites than to CU facilities. In the \hydrogenscenario{}, where all carbon for CU and CS must come from DAC, the hydrogen transport topology remains almost identical to the net-zero scenarios (for detailed cost contributions see sections~\ref{sec:cost_comparison} and~\ref{sec:operation_nn} in the Appendix).

Overall, this leads to the net cost increases of 78~bn€/a in the \carbonscenario{}, by 80~bn€/a in the \hydrogenscenario{}, and by 78~bn€/a in the \hybridscenario{}. The resulting cost-benefit of the \hybridscenario{} in comparison to the other scenarios is 6~bn€/a.

@@

\subsection{Limitations}
\label{sec:limitations}
Despite the comprehensive nature of the model, some limitations affect the validity of the results, as well as important discussion points that need to be put into context.
In reality, energy demand and renewable supply can only be estimated for the next few hours/days. The model's perfect foresight may lead to non-reproducible technology dispatch which neglects uncertainties from long-term predictions and necessary reserve margins.
Demand and emissions from industrial clusters heavily drive the modeling results. Allowing the model to relocate these, incorporate flexibility measures or alter underlying processes, may lead to less dependency on \carbon{} and \hydrogen{} networks.
The transport of fuel (FT, gas, oil) between the regions is not limited which may overestimate their flexibility. However, the distributed CU systems, as enabled through the hydrogen network, align well with today's distributed locations for oil refineries, providing kerosene for aviation and naphtha for industry. The assumption to run Fischer-Tropsch facilities on baseload with at least 90\% of their nominal capacity is likely shaping the deployment of CU, underestimating potential important flexibilities.
The assumed cost projections on technologies are subject to uncertainties. As shown in \ref{sec:subsidy}, the effect of a 50\% cost reduction on \carbon{} pipelines alters total system costs to a small extent (below 1\%). However, a different set of costs on DAC and electrolysis may well impact the model results.
Considering different strategies for carbon removal, DAC deployment could be in parts replaced if larger sustainable biomass sources are assumed, as shown in~\cite{lauerCrucialRoleBioenergy2023}.

Finally, our energy system model assumes total autarky of the European system and neglects potential imports from elsewhere. Depending on the type of energy carrier and whether it is \carbon{} neutral or not, allowing imports can change the optimal design and operation of the system quite significantly, as discussed in~\cite{neumannEnergyImportsInfrastructure2024}. In our model, we see that when higher sequestration rates are allowed, the model imports more fossil oil, which reduces the use of CU in Central Europe and thus the hydrogen transport from South to Central Europe (see section~\ref{sec:higher_sequestration}). While this substitution reduces the need for Fischer-Tropsch fuel production, it also requires significant carbon sequestration within Europe to offset the emissions from fossil oil use. If the imported oil were carbon neutral or 'green', there would be no need for dedicated sequestration infrastructure as the emissions would already be accounted for. In such a case, the overall energy infrastructure would remain largely the same, minus the additional carbon sequestration capacity needed to offset the emissions from fossil imports. Thus both the \hydrogen{} and \carbon{} networks would be less relevant.
In the context of green hydrogen imports, as highlighted by~\cite{neumannEnergyImportsInfrastructure2024}, the model is expected to reduce the use of electrolysis and methanolisation in southern European countries, as well as hydrogen transport routes from there to central Europe. However, the deployment of CU near point sources of \carbon{} would still be as in our scenario. The same applies to the deployment of a carbon network driven by fixed sequestration sites.
